Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-03 fib2.tex
resolved bibkey conflict for new jw00t ref

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014, Wootton2015}. Nonabelian anyon models are especially interesting 
because braiding and fusion of these anyons in general allows for the implementation of universal quantum 
computation. However, the initial studies of error-correction in nonabelian anyon systems have focused on specific models, such as the Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal. The general dynamics of these particular anyon models is known to be efficiently classically simulable, a fact
that was exploited to enable efficient simulation of error correction in these systems. When considering more general anyon models, their ability to perform universal quantum computation would seem a significant barrier to their simulation on a classical computer. While simulation of general dynamics does indeed seem intractable, we argue that the kinds of processes that are typical of thermal noise are sufficiently structured to allow for their classical simulation in the regimes where we expect successful error correction to be possible. This insight allows us to simulate the noise and recovery processes for a quantum code based on a universal anyon model.

@@

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, producing a \emph{syndrome} of occupied sites. Following this, it forms clusters of nearby anyons and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when there is at most a single cluster remaining (see \Fref{f:decode}).
Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or when it performs a homologically nontrivial operation, thereby altering the codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

>>>>

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014, Wootton2015b}. Nonabelian anyon models are especially interesting 
because braiding and fusion of these anyons in general allows for the implementation of universal quantum 
computation. However, the initial studies of error-correction in nonabelian anyon systems have focused on specific models, such as the Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal. The general dynamics of these particular anyon models is known to be efficiently classically simulable, a fact
that was exploited to enable efficient simulation of error correction in these systems. When considering more general anyon models, their ability to perform universal quantum computation would seem a significant barrier to their simulation on a classical computer. While simulation of general dynamics does indeed seem intractable, we argue that the kinds of processes that are typical of thermal noise are sufficiently structured to allow for their classical simulation in the regimes where we expect successful error correction to be possible. This insight allows us to simulate the noise and recovery processes for a quantum code based on a universal anyon model.

@@

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015b},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, producing a \emph{syndrome} of occupied sites. Following this, it forms clusters of nearby anyons and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when there is at most a single cluster remaining (see \Fref{f:decode}).
Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or when it performs a homologically nontrivial operation, thereby altering the codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.
